---
title: "SS Calculations for Transport Studies"
author: "Whitney Knauer"
date: "today"
output: html_document
---

```{r}
#| echo: false
#| warning: false

library(tidyverse)
library(skimr)
library(dplyr)
library(ggplot2)
library(ggbeeswarm)
library(lme4)
library(car)
library(emmeans)
library(gmodels)
library(epiDisplay)
library(psych)
library(doBy)
library(ggpubr)
library(gridExtra)
library(simr)
library(pwr)
library(broom)
library(gt)
library(showtext)
library(TOSTER)
library(epiR)
```

## Sample Size calculation for mortality difference.

RV all cause mortality is 1.6% from arrival to 60d. Which seems too low, wonder what the denominator is?

```{r}
# Function to calculate sample size for a difference in proportions
sample_size_proportions <- function(p1, p2, power = 0.8, alpha = 0.05) {
  result <- power.prop.test(p1 = p1, p2 = p2, power = power, sig.level = alpha)
  return(ceiling(result$n)) # Round up to ensure sufficient sample size
}

# Example usage:
p1 <- 0.016  # Proportion in group 1
p2 <- 0.032  # Proportion in group 2
sample_size <- sample_size_proportions(p1, p2)
cat("Required sample size per group:", sample_size, "\n")

sample_size*1.25
```

If a treatment doubles the mortality rate from 1.6% to 3.2%, 1436 calves are needed per group. Need to also account for trailer load (clustering), source farm, etc. Will inflate samples size by 25%, so new number needed per group \~1800.

## Sample size calculation for risk difference (injury)

```{r}

# Function to calculate sample size for a risk difference
sample_size_risk_difference <- function(p1, p2, power = 0.8, alpha = 0.05) {
  result <- power.prop.test(p1 = p1, p2 = p2, power = power, sig.level = alpha)
  return(ceiling(result$n)) # Round up to ensure sufficient sample size
}


p1 <- 0.01  # Risk in high/normal stocking density
p2 <- 0.02  # Risk in low stocking density group

sample_size <- sample_size_risk_difference(p1, p2)
cat("Required sample size per group:", sample_size, "\n")

sample_size*1.25
```

To find a (p2-p1) risk difference or a risk difference of 1% (risk of injury increases with lower stocking density), a sample size of 2391 is needed. To inflate for clustering (trailer, source farm, etc) we will increase our ss by 25% and thus will require 2900 calves per group.

## Sample size calculation for relative risk of injury

```{r}
# Function to calculate sample size for risk ratio using power.prop.test
sample_size_risk_ratio <- function(p1, RR, power = 0.8, alpha = 0.05) {
  p2 <- p1 * RR  # Compute expected proportion in the exposed group
  result <- power.prop.test(p1 = p1, p2 = p2, power = power, sig.level = alpha)
  return(ceiling(result$n)) # Round up to the nearest whole number
}

# Example usage:
p1 <- 0.01  # Risk in high/normal stocking rate group
RR <- 2    # Expected risk ratio, risk is 2X in low stocking rate group
sample_size <- sample_size_risk_ratio(p1, RR)
cat("Required sample size per group:", sample_size, "\n")


```

Same sample size as the RD ss because really saying the same thing with these two sets of arguments.

## Sample Size calculation for scours difference in SD study.3 treatment groups 3ft/3.5ft/4.5ft

RV scour rate is 12% from arrival to 60d.

```{r}
# Function to calculate sample size for a difference in proportions
sample_size_proportions <- function(p1, p2, power = 0.8, alpha = 0.016) {
  result <- power.prop.test(p1 = p1, p2 = p2, power = power, sig.level = alpha)
  return(ceiling(result$n)) # Round up to ensure sufficient sample size
}

# Example usage:
p1 <- 0.30  # Proportion in group 1
p2 <- 0.26  # Proportion in group 2
sample_size <- sample_size_proportions(p1, p2)
cat("Required sample size per group:", sample_size, "\n")

sample_size*1.25
```
